---
title: "R Notebook"
output: html_notebook
---

```{bash}
library('tidyverse')
```

Viewing the original data

```{r}
passengers <- read.csv('spaceship-titantic-data.csv')
view(passengers)
```

**Variables to wrangle:**

-   age -\> age groups

-   cabins -\> deck, num, side

```{r}
# Age Groups
get_age_group <- function(pass_age){
  group <- NA
  if(is.na(pass_age)){group <- NA}
  else{
    age <- as.integer(pass_age)
    if(age < 5){group <- "Infant"}
    else if(age < 18){group <- "Adolescent"}
    else if(age < 35){group <- "Young Adult"}
    else if(age < 60){group <- "Adult"}
    else{group <- "Older Adult"}
  }
  return(group)
}

passengers$age_group <- unlist(lapply(passengers[["Age"]], get_age_group))
head(passengers)
```

```{r}
# Cabin Components
get_deck <- function(cabin){
  if(is.na(cabin) || cabin == ""){deck <- NA}
  else{
    c_vector <- unlist(strsplit(cabin, "/"))
    deck <- c_vector[1]
  }
  return(deck)
}
get_num <- function(cabin){
  if(is.na(cabin) || cabin == ""){num <- NA}
  else{
    c_vector <- unlist(strsplit(cabin, "/"))
    num <- c_vector[2]
  }
  return(num)
}
get_side <- function(cabin){
  if(is.na(cabin) || cabin == ""){side <- NA}
  else{
    c_vector <- unlist(strsplit(cabin, "/"))
    side <- c_vector[3]
  }
  return(side)
}

passengers$cabin_deck <- unlist(lapply(passengers[["Cabin"]], get_deck))
passengers$cabin_num <- unlist(lapply(passengers[["Cabin"]], get_num))
passengers$cabin_side <- unlist(lapply(passengers[["Cabin"]], get_side))

head(passengers)
```

Looking into the data...

```{r}
head(passengers)
```

**General Passenger Demographics**

-   Age/Age Groups

```{r}
age_counts <- passengers %>% count(Age)
age_counts %>% drop_na() %>% filter(Age > 0)
age_counts$group <- unlist(lapply(age_counts[["Age"]], get_age_group))

age_group_counts <- passengers %>% count(age_group)
age_group_counts %>% drop_na()

ppt_colors <- c("#512887", "#AE0c00", "#062A79", "#CB0815", "#891446")
ggplot(data = age_counts) + 
  geom_col(mapping = aes(x = Age, y = n, fill = group)) +
  scale_fill_manual(values= ppt_colors)
```

-   Home Planets and Destinations

```{r}
planet_counts <- passengers %>% count(HomePlanet)
planet_counts <- planet_counts %>% drop_na() %>% filter(HomePlanet != "")
destination_counts <- passengers %>% count(Destination)
destination_counts <- destination_counts %>% drop_na() %>% filter(Destination != "")

ggplot(planet_counts, aes(x="", y=n, fill=HomePlanet)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_manual(values= c(ppt_colors[1], ppt_colors[5], "#08369B"))+ 
  theme_void()

ggplot(destination_counts, aes(x="", y=n, fill=Destination)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_manual(values= c(ppt_colors[4], ppt_colors[3], "#631A87"))+ 
  theme_void()
```

-   CryoSleep

```{r}
sleep_counts <- passengers %>% count(CryoSleep)
sleep_counts <- sleep_counts %>% drop_na() %>% filter(CryoSleep != "")

ggplot(data=sleep_counts, aes(x=CryoSleep, y=n, fill = CryoSleep)) +
  geom_bar(stat="identity") + 
  scale_fill_manual(values= c("#fd6ab9", "#6a6ffd"))

sleep_counts

```

-   Cabin Breakdowns

```{r}
deck_count <- passengers %>% count(cabin_deck)
num_count <- passengers %>% count(cabin_num)
side_count <- passengers %>% count(cabin_side)

cabin_density <- passengers %>% group_by(cabin_side, cabin_deck) %>% count(cabin_deck) %>% drop_na()


ggplot(data = cabin_density) + 
  geom_tile(mapping = aes(x = cabin_side, y = cabin_deck, fill = n)) +
  scale_fill_gradient(low="blue", high="red")

```

-   VIP

```{r}
vip_counts <- passengers %>% count(VIP)
vip_counts <- vip_counts %>% drop_na() %>% filter(VIP != "")

ggplot(data=vip_counts, aes(x=VIP, y=n, fill = VIP)) +
  geom_bar(stat="identity") + 
  scale_fill_manual(values= c("#ae6afd", "#f242f8"))
vip_counts

```

-   Transported

```{r}
transport_count <- passengers %>% count(Transported)

ggplot(data=transport_count, aes(x=Transported, y=n, fill = Transported)) +
  geom_bar(stat="identity") + 
  scale_fill_manual(values= c("#ce0e74", "#0e14ce"))

transport_count

```

Take aways:

**Transported Passenger Demographics**

Making Transported Passengers a separate data frame to work with easier

```{r}
transported_passengers <- passengers %>% filter(Transported == "True")
transported_passengers

```

-   Age/Age Groups

```{r}
t_age_counts <- transported_passengers %>% count(Age)
t_age_counts %>% drop_na() %>% filter(Age > 0)
t_age_counts$group <- unlist(lapply(t_age_counts[["Age"]], get_age_group))

t_age_group_counts <- transported_passengers %>% count(age_group)
t_age_group_counts %>% drop_na()

ppt_colors <- c("#512887", "#AE0c00", "#062A79", "#CB0815", "#891446")
ggplot(data = t_age_counts) + 
  geom_col(mapping = aes(x = Age, y = n, fill = group)) +
  scale_fill_manual(values= ppt_colors)
```

-   Home Planets and Destinations

```{r}
t_planet_counts <- transported_passengers %>% count(HomePlanet)
t_planet_counts <- t_planet_counts %>% drop_na() %>% filter(HomePlanet != "")
t_destination_counts <- transported_passengers %>% count(Destination)
t_destination_counts <- t_destination_counts %>% drop_na() %>% filter(Destination != "")

ggplot(t_planet_counts, aes(x="", y=n, fill=HomePlanet)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_manual(values= c(ppt_colors[1], ppt_colors[5], "#08369B"))+ 
  theme_void()

ggplot(t_destination_counts, aes(x="", y=n, fill=Destination)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_manual(values= c(ppt_colors[4], ppt_colors[3], "#631A87"))+ 
  theme_void()
```

-   CryoSleep

```{r}
t_sleep_counts <- transported_passengers %>% count(CryoSleep)
t_sleep_counts <- t_sleep_counts %>% drop_na() %>% filter(CryoSleep != "")

ggplot(data=t_sleep_counts, aes(x=CryoSleep, y=n, fill = CryoSleep)) +
  geom_bar(stat="identity") + 
  scale_fill_manual(values= c("#fd6ab9", "#6a6ffd"))
```

-   Cabin Breakdowns

```{r}
t_deck_count <- transported_passengers %>% count(cabin_deck)
t_num_count <- transported_passengers %>% count(cabin_num)
t_side_count <- transported_passengers %>% count(cabin_side)

t_cabin_density <- transported_passengers %>% group_by(cabin_side, cabin_deck) %>% count(cabin_deck) %>% drop_na()


ggplot(data = t_cabin_density) + 
  geom_tile(mapping = aes(x = cabin_side, y = cabin_deck, fill = n)) +
  scale_fill_gradient(low="blue", high="red")
```

-   VIP

```{r}
t_vip_counts <- transported_passengers %>% count(VIP)
t_vip_counts <- t_vip_counts %>% drop_na() %>% filter(VIP != "")

ggplot(data=t_vip_counts, aes(x=VIP, y=n, fill = VIP)) +
  geom_bar(stat="identity") + 
  scale_fill_manual(values= c("#ae6afd", "#f242f8"))
```

**Transported Passenger Demographics v. General Passenger Demographics**

-   Age/Age Groups

```{r}
comp_age_group <- passengers %>% group_by(age_group, Transported) %>% 
  count() %>% drop_na()


comp_age_group
ggplot(data=comp_age_group, aes(x=age_group, y=n, fill = Transported)) +
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_manual(values= c("red", "blue"))
```

-   Home Planets and Destinations

```{r}
par(mfrow=c(1,2))
non_t_hd_density <- passengers %>% group_by(HomePlanet, Destination) %>% 
  filter(Transported == "False") %>% 
  count() %>% drop_na() %>% 
  filter(HomePlanet != "") %>% 
  filter(Destination != "")
non_t_hd_density
t_hd_density <- passengers %>% group_by(HomePlanet, Destination) %>% 
  filter(Transported == "True") %>% 
  count() %>% drop_na() %>% 
  filter(HomePlanet != "") %>% 
  filter(Destination != "")
t_hd_density

ggplot(data = non_t_hd_density) + 
  geom_tile(mapping = aes(x = HomePlanet, y = Destination, fill = n)) +
  scale_fill_gradient(low="blue", high="red") + 
  ggtitle("Non-Transported Passengers Home Planet v Destination")
ggplot(data = t_hd_density) + 
  geom_tile(mapping = aes(x = HomePlanet, y = Destination, fill = n)) +
  scale_fill_gradient(low="blue", high="red") + 
  ggtitle("Transported Passengers Home Planet v Destination")
```

-   CryoSleep

```{r}
#bar graph
comp_cryo_sleep <- passengers %>% group_by(CryoSleep, Transported) %>% 
  count() %>% drop_na() %>% filter(CryoSleep != "")


comp_cryo_sleep

ggplot(data=comp_cryo_sleep, aes(x=CryoSleep, y=n, fill = Transported)) +
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_manual(values= c("red", "blue"))
```

-   Cabin Breakdowns

```{r}
# side by side density maps
par(mfrow=c(1,2))
non_t_c_density <- passengers %>% group_by(cabin_side, cabin_deck) %>% 
  filter(Transported == "False") %>% 
  count() %>% drop_na() %>% 
  filter(cabin_side != "") %>% 
  filter(cabin_deck != "")
non_t_c_density
t_c_density <- passengers %>% group_by(cabin_side, cabin_deck) %>% 
  filter(Transported == "True") %>% 
  count() %>% drop_na() %>% 
  filter(cabin_side != "") %>% 
  filter(cabin_deck != "")
t_c_density

ggplot(data = non_t_c_density) + 
  geom_tile(mapping = aes(x = cabin_side, y = cabin_deck, fill = n)) +
  scale_fill_gradient(low="blue", high="red") + 
  ggtitle("Non-Transported Passengers Cabin Location")
ggplot(data = t_c_density) + 
  geom_tile(mapping = aes(x = cabin_side, y = cabin_deck, fill = n)) +
  scale_fill_gradient(low="blue", high="red") + 
  ggtitle("Transported Passengers Cabin Location")
```

-   VIP

```{r}
#bar graph

comp_vip <- passengers %>% group_by(VIP, Transported) %>% 
  count() %>% drop_na() %>% filter(VIP != '')


comp_vip
ggplot(data=comp_vip, aes(x=VIP, y=n, fill = Transported)) +
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_manual(values= c("red", "blue"))
```

```{r}
cc_density <- passengers %>% group_by(cabin_side, cabin_deck) %>% 
  filter(Transported == "True") %>% 
  filter(CryoSleep == "True") %>% 
  count() %>% drop_na() %>% 
  filter(cabin_side != "") %>% 
  filter(cabin_deck != "")

ggplot(data = cc_density) + 
  geom_tile(mapping = aes(x = cabin_side, y = cabin_deck, fill = n)) +
  scale_fill_gradient(low="blue", high="red") + 
  ggtitle("Transported Passengers in Cryo Sleep Cabin Location")


ncc_density <- passengers %>% group_by(cabin_side, cabin_deck) %>% 
  filter(Transported == "True") %>% 
  filter(CryoSleep == "False") %>% 
  count() %>% drop_na() %>% 
  filter(cabin_side != "") %>% 
  filter(cabin_deck != "")

ggplot(data = ncc_density) + 
  geom_tile(mapping = aes(x = cabin_side, y = cabin_deck, fill = n)) +
  scale_fill_gradient(low="blue", high="red") + 
  ggtitle("Transported Passengers not in Cryo Sleep Cabin Location")
```
